Automated Assessment of Residual Plots with Computer Vision Models

### 2025-06-05 paper.Rmd
Fix typo
Update section 2.2

Taking into account the implementation cost and the need for model interpretability, we used the single residual plot input format in this chapter.

@@

Given that the input is a single residual plot represented as a fixed-resolution image, we can choose the output from the computer vision model to be either binary (classification) or numeric (regression).

@@

@@

In this study, we chose to define and use a distance between a true residual plot and a theoretically "good" residual plot. This is further explained in Section \ref{sec-model-distance-between-residual-plots}. @vo2016localizing have also demonstrated that defining a proper distance between images can enhance the matching accuracy in image search compared to a binary outcome model.

>>>>

Taking into account the implementation cost and the need for model interpretability, we used the single residual plot input format in this paper.

@@

Given that the input is a single residual plot represented as a fixed-resolution image, the output of the computer vision model can be designed for various tasks: binary classification, categorical classification with multiple classes, or numeric regression.

@@

For a multi-class outcome, the model could classify images into categories such as "contains outliers" vs. "does not contain outliers" or "is non-linear" vs. "not non-linear," among other diagnostic features. While this approach is appealing in theory, it essentially reverts to the traditional testing framework, where tests focus on detecting specific types of model violations in isolation.

One significant advantage of the lineup protocol for residual diagnostics is its ability to detect multiple model violations simultaneously, provided these violations produce visually distinguishable differences in the plots. In contrast, training a classification model would require explicitly defining and mapping all potential violations of interest, or at least the most common ones, which can be impractical. If the goal is to identify specific cases, conventional statistical tests grounded in regression model assumptions are often more effective. Computer vision models may be less suited to such tasks.

@@

In this study, we chose to define and use a distance between a true residual plot and a theoretically "good" residual plot. This is further explained in @sec-model-distance-between-residual-plots. @vo2016localizing have also demonstrated that defining a proper distance between images can enhance the matching accuracy in image search compared to a binary outcome model.
